---
title: "1.4 — Perfect Competition II — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Thursday, February 2, 2023*
:::

<!--::: callout-warning
# Upcoming Assignment
Please complete the [preliminary survey](/assignments/00-preliminary-survey.qmd) by next class.
:::-->

## <i class="fas fa-info-circle fa-lg"></i> Overview

We continue our review/extension of perfect competition by looking at how to derive the equilibrium for the industry over the long run, where entry and exit are possible. We also discuss the role of economic rents.

## <i class="fas fa-book-reader fa-lg"></i> Readings

### Recommended Readings

- <i class="fas fas fa-book"></i> Ch. 2.1-2.2, 3.1 in [Church & Ware, 2000](https://works.bepress.com/jeffrey_church/23/)

## <i class="fas fa-dumbbell"></i> Practice

Today you will be working on practice problems:

- [<i class="fas fa-file-pdf"></i> Practice Problems](../files/practice/1.4-practice-pdf.pdf)
- [<i class="fas fa-external-link-square-alt"></i> Answer Key](../files/practice/1.4-practice-answers.html)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("1.4-slides")
```

<!--## <i class="fas fa-laptop-code"></i> Assignments

### Preliminary Math Survey Due Next Class

Please take the [preliminary survey](/assignment/00-preliminary-survey) on your math background by next class. This will help us all have a productive semester together.
-->