---
title: "1.6 — Monopoly II — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Thursday, February 9, 2023*
:::

<!--::: callout-warning
# Upcoming Assignment
Please complete the [preliminary survey](/assignments/00-preliminary-survey.qmd) by next class.
:::-->

## <i class="fas fa-info-circle fa-lg"></i> Overview

We continue looking at monopoly and see what are the economic consequences of monopoly.

## <i class="fas fa-book-reader fa-lg"></i> Readings

### Recommended Readings

This is a part of our review of major production concepts from Microeconomics. You can consider looking at the following in the textbook if you are rusty or want to know more about these concepts:   

- <i class="fas fas fa-book"></i> Ch. 2.1-2.2, 3.1, 4.1—4.2, 4.4-4.5 in [Church & Ware, 2000](https://works.bepress.com/jeffrey_church/23/)

- [<i class="fas fa-file-pdf"></i> Harberger (1954), "Monopoly and Resource Allocation"](/readings/Harberger-1954.pdf)
- [<i class="fas fa-file-pdf"></i> Tullock (1967), "The Welfare Cost of Tariffs, Monopolies, and Theft"](/readings/Tullock-1967.pdf)
- [<i class="fas fa-file-pdf"></i> Mitchell (2012), "The Pathology of Privilege"](/readings/Mitchell-2012.pdf)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("1.6-slides")
```