---
title: "2.2 — Cournot Competition — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Thursday, February 23, 2023*
:::

::: callout-warning
# Upcoming Assignment
[Problem Set 1](/assignment/01-problem-set) (1.1-1.8) is due by 11:59 PM Thursday February 23 on Blackboard Assignments.
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we cover our first, and perhaps the most famous and "base" model of Oligopoly: Cournot's model of firms *simultaneously* competing over *quantity.*

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fas fa-book"></i> Ch. 8.1-8.2, 7 in [Church & Ware, 2000](https://works.bepress.com/jeffrey_church/23/)

## <i class="fas fa-dumbbell"></i> Practice

Today you will be working on practice problems:

- [<i class="fas fa-file-pdf"></i> Practice Problems](../files/practice/2.2-practice-pdf.pdf)
- [<i class="fas fa-file-pdf"></i> Answer Key](../files/practice/2.2-practice-answers.html)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("2.2-slides")
```

## <i class="fas fa-laptop-code"></i> Assignments

### Problem Set 1 Due Thursday February 23

[Problem Set 1](/assignment/01-problem-set) (1.1-1.8) is due by 11:59 PM Thursday February 23 on Blackboard Assignments.
