---
title: "3.2 — Stackelberg Competition — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Tuesday, March 7, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we extend our understanding of game theory to talk about how collusion can be sustainable.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fas fa-book"></i> Ch. 7, 9, 10 in [Church & Ware, 2000](https://works.bepress.com/jeffrey_church/23/)

### Recommended/Referenced Readings

- [<i class="fas fa-file-pdf"></i> Stigler, 1964, “A Theory of Oligopoly”](/files/readings/Stigler-1964.pdf)

Chapter 7 is a good overview of static game theory - read it if you have not yet, or want some more of the basics. We will focus simply on Nash equilibrium and the possibility of multiple Nash equilibria.

Chapter 9 is a good overview of sequential games and repeated games, which we will introduce today.

Chapter 10 discusses oligopoly in light of how collusion can be sustained in dynamic settings.

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("3.2-slides")
```