---
title: "3.3 — Entry Deterrence and Commitment — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Tuesday, March 21, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we discuss entry deterrence, entry games, subgame perfection, and commitments in game theory.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fas fa-book"></i> Ch. 9, 13, 14 in [Church & Ware, 2000](https://works.bepress.com/jeffrey_church/23/)

Chapter 9 covers sequential games, games in extensive form, subgames, and subgame perfection.

Chapter 13 covers strategic behavior, limit output/pricing, and entry deterrence.

Chapter 14 gives more detailed models of entry deterrence and contestable markets.

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("3.3-slides")
```