---
title: "3.3 — Entry Deterrence and Commitment — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Tuesday, March 21, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we start looking at what is known in economics as "the theory of the firm." We begin with asking the question, "if markets are so great, why are there firms?" We look at Adam Smith's idea of the division of labor, and when specialization makes sense. Then we look at Coase's key contribution - that firms exist because there are transaction costs of using markets. We wrap up with Stigler's relation of the division of labor to the life cycle of industries.

For the next few classes, we will look at other aspects of the theory of the firm, and address why firms look and behave the way they do, by understanding the economic problems that organizations try to solve.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- [<i class="fas fa-file-pdf"></i> Coase (1937), "The Nature of the Firm"](/files/readings/Coase-1937.pdf)
- [<i class="fas fa-external-link-square-alt"></i> Munger (2007), "I'll Stick With These: Some Sharp Observations on the Division of Labor"](https://www.econlib.org/library/Columns/y2007/Mungerpins.html)
- [<i class="fas fa-external-link-square-alt"></i> Smith (1776), *Wealth of Nations*, Book I, Chs. 1-3](https://www.econlib.org/library/Smith/smWN.html)
- [<i class="fas fa-file-pdf"></i> Stigler (1951), "The Division of Labor is Limited by the Extent of the Market"](/files/readings/Stigler-1951.pdf)

The free IO textbook does a good job covering the theory of the firm in Chapter 3:

- <i class="fas fas fa-book"></i> Ch. 3.2 in [Church and Ware](https://works.bepress.com/jeffrey_church/23/), 2000

Much of the rest of their discussion in Chapter 3 we will cover over the next few classes (asset specificity theory, vertical integration, etc.)

### Discussion Board 

Starting next week, we will occasionally have a weekly discussion board open on Blackboard. (Not every week. I will let you know when we have one open at the beginning of each week.) You will be expected to contribute to the discussion board at least twice by Sunday night. Your weekly contribution will be graded out of 5 points. At the end of the semester, I will apply the average of your weekly participation grades to apply (20%) towards your final course grade.

I am interested in your thoughts, reactions, comments, and questions about any of the material (lectures and/or readings). You do not need to write more than a paragraph. Anything more than that, including continuing to reply to each others’ thoughts, questions, or comments, (which I strongly hope you do!) is solely based on your own interest and curiosity. I will jump in to answer questions the group is stuck on, give my two cents, and stir the pot as needed. I strontly hope we still keep a conversation going and can learn from each other, that was always my goal, not to lecture at you!

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("4.1-slides")
```