---
title: "4.2 — The Firm as Nexus of Contracts — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Thursday, March 30, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we continue looking at the "theory of the firm" launched by Coase (1937). We explore a famous interpretation (among other intepretations) of the firm as a "**nexus of contracts**" in the marketplace. This comes from a famous paper by **Alchian and Demsetz (1972)**, which views the firm not as a command-and-control hierarchy, but another nexus of voluntary market transactions, where collaborating factors of production find it in their interest to hire a monitor to minimize shirking, and give the monitor residual claimancy over the output.

We extend this view by discussing **agency costs**, which often (but not exclusively) arise from the separation of ownership and control, as in corporate businesses. We discuss the famous paper by **Jensen and Meckling (1976)** on agency costs, and then discuss **corporate governance** strategies on how firms attempt to minimize these agency costs, including a final discussion of the **market for corporate control.**

## <i class="fas fa-book-reader fa-lg"></i> Readings

### Recommended Readings:

- [<i class="fas fa-file-pdf"></i> Alchian and Demsetz, 1972, "Production, Information Costs, and Economic Organization"](/files/readings/Alchian-Demsetz-1972.pdf)
- [<i class="fas fa-file-pdf"></i> Jensen and Meckling, 1976, "Theory of the Firm: Managerial Behavior, Agency Costs and Ownership Structure"](/files/readings/Jensen-Meckling-1976.pdf)
- [<i class="fas fa-file-pdf"></i> Jensen and Murphy, 1990, "CEO Incentives - It's Not How Much You Pay, But How"](/files/readings/Shleifer-Vishny-1997.pdf)
- [<i class="fas fa-file-pdf"></i> Manne, 1965, "Mergers and The Market for Corporate Control"](/files/readings/Manne-1965.pdf)
- [<i class="fas fa-file-pdf"></i> Murphy, 2012, "Executive Compensation: Where We Are, and How We Got There"](/files/readings/Murphy-2012.pdf)
- [<i class="fas fa-file-pdf"></i> Shleifer and Vishney, 1997, "A Survey of Corporate Governance"](/files/readings/Shleifer-Vishny-1997.pdf)

Alchian and Demsetz (1972) is the most important to read (and the shortest and most readable), followed by Jensen and Meckling (1976), which, after the first section, gets too lengthy and technical.

The free IO textbook only very briefly covers Alchian and Demsetz (1972) and nexus of contracts theory. It then launches into a very technical example. This book is less useful today than for last class and for next class:

- <i class="fas fas fa-book"></i> pp.79-81 in [Church and Ware](https://works.bepress.com/jeffrey_church/23/)

### Discussion Board 

Starting next week, we will occasionally have a weekly discussion board open on Blackboard. (Not every week. I will let you know when we have one open at the beginning of each week.) You will be expected to contribute to the discussion board at least twice by Sunday night. Your weekly contribution will be graded out of 5 points. At the end of the semester, I will apply the average of your weekly participation grades to apply (20%) towards your final course grade.

I am interested in your thoughts, reactions, comments, and questions about any of the material (lectures and/or readings). You do not need to write more than a paragraph. Anything more than that, including continuing to reply to each others’ thoughts, questions, or comments, (which I strongly hope you do!) is solely based on your own interest and curiosity. I will jump in to answer questions the group is stuck on, give my two cents, and stir the pot as needed. I strontly hope we still keep a conversation going and can learn from each other, that was always my goal, not to lecture at you!

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("4.2-slides")
```