---
title: "4.3 — Asset Specificity & Vertical Integration — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Tuesday, April 4, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we cover another major framework for studying the theory of the firm, the theory of asset specificity. This approach focuses on the threat of opportunism by parties that have signed a contract and made specific investments (sunk costs) by depending on one another: parties could try to extort greater quasi-rents from one another, or else threaten to pull out of the contract, harming the other party.

Better-written enforceable contracts are one method of trying to minimize this type of transaction cost, but another method is vertical integration - merging the two parties (e.g. perhaps a manufacturer and a retailer) as a single firm, that does not have to worry about such opportunism.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fas fa-book"></i> pp.69-79 in [Church and Ware](https://works.bepress.com/jeffrey_church/23/)

The free IO textbook pretty extensively covers asset specificity theory, appropriable quasi-rents, and even the Fisher Body example (it's very famous):

### Recommended Readings:

- [<i class="fas fa-file-pdf"></i> Klein, Crawford, and Alchian, 1978, "Vertical Integration, Appropriable Rents, and the Competitive Contracting Process"](/files/readings/Klein-Crawford-Alchian-1978.pdf)
- [<i class="fas fa-file-pdf"></i> Williamson, 1983, "Credible Commitments: Using Hostages to Support Exchange"](/files/readings/Williamson-1983.pdf)
- [<i class="fas fa-file-pdf"></i> Williamson, 1998, "The Institutions of Governance"](/files/readings/Williamson-1998.pdf)
- [<i class="fas fa-file-pdf"></i> Williamson, 2009, "Transaction Cost Economics: The Natural Progression" (Nobel Prize Lecture)](https://www.nobelprize.org/prizes/economic-sciences/2009/williamson/lecture/)

### Discussion Board 

Starting this week, we will occasionally have a weekly discussion board open on Blackboard. (Not every week. I will let you know when we have one open at the beginning of each week.) You will be expected to contribute to the discussion board at least twice by Sunday night. Your weekly contribution will be graded out of 5 points. At the end of the semester, I will apply the average of your weekly participation grades to apply (20%) towards your final course grade.

I am interested in your thoughts, reactions, comments, and questions about any of the material (lectures and/or readings). You do not need to write more than a paragraph. Anything more than that, including continuing to reply to each others’ thoughts, questions, or comments, (which I strongly hope you do!) is solely based on your own interest and curiosity. I will jump in to answer questions the group is stuck on, give my two cents, and stir the pot as needed. I strontly hope we still keep a conversation going and can learn from each other, that was always my goal, not to lecture at you!

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("4.3-slides")
```