---
title: "4.5 — Entrepreneurship, Uncertainty, and Market Process — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Tuesday, April 11, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we wrap up Unit 4 and try to accomplish several things before we move into empirically trying to measure and regulate market power in our final Unit.

First, we consider the role of entrepreneurship in the theory of the firm.

Second, we consider the role of uncertainty in our models of markets. 

The combination of these ideas leads to some stark realizations about the way we model the real world, and its implications for understanding things like what “competition” actually means, and how to attain it, and such, what regulation should be. 

All too often, people learn the models of basic economics (i.e. utility & profit maximization, market equilibrium, perfect competition) and make one of the following mistakes:

1. They see the real world looking nothing like the model and conclude the model is *useless.*
2. They see the real world looking nothing like the model, and conclude the real world is *wrong* and must be *fixed* to work like the model works.
3. They believe the model to be a 100% *accurate* description of the real world.

The essence of this lecture is very subtle, but very important. Hence, I have tried to give you some well-known readings.

Knight’s (1920) book is the most famous discussion of the perfect competition model, and how real world competition differs from the model because of uncertainty, which gives rise to the possibility of positive profit in the real world.

Hayek (1945) is one of the most famous papers in economics, arguing how the economy is not a constrained optimization problem, and that central planning faces an enormous knowledge problem, as prices are needed to tap into the fragmented, tacit knowledge dispersed across individuals (that no single person or entity can know) to coordinate the behavior of millions of people.

Alchian (1950) argues that the behavioral assumption of firms maximizing profits is incorrect, because there is no way anything can be maximized in a world of uncertainty (see Knight). It is better to view market competition in evolutionary terms, firms survive simply by earning positive profits (rather than maximum profits) by best adapting to the competitive environment — whether by intention and skill, or by sheer luck: “Even in a world of stupid men there would be profits.”

The video lecture by Gerd Gigerenzer provides a more modern approach to understanding how to model the world under uncertainty: that it is impossible to optimize, but instead to use heuristics. This stands as a strong criticism of behavioral economics (which focuses on the supposed irrationality of decision-making).

Similarly, Smith’s (2003) Nobel Prize article discusses this.

## <i class="fas fa-book-reader fa-lg"></i> Readings

### Recommended Readings:

- [<i class="fas fa-file-pdf"></i> Alchian, Armen A, 1950, "Uncertainty, Evolution, and Economic Theory"](/files/readings/Alchian-1950.pdf)
- <i class="fab fa-youtube"></i> Gigerenzer, Gerd, 2012, ["What Can Economists Know?"](https://www.youtube.com/watch?v=DdEEwoKkfMA)], Institute for New Economic Thinking's (INET) Paradigm Lost Conference, Berlin
- <i class="fas fas fa-book"></i> Hayek, Friedrich A, 1948, ["The Meaning of Competition"](https://mises.org/library/meaning-competition), Chapter V in *Individualism and Economic Order*
- <i class="fas fas fa-book"></i> Kirzner, Israel M, 1973, [*Competition and Entrepreneurship*](http://www.sjsu.edu/people/john.estill/courses/158-s15/Israel%20Kirzner%20-%20Competition%20And%20Entrepreneurship.pdf)
- <i class="fas fas fa-book"></i> Knight, Frank H, 1921, [*Risk, Uncertainty, and Profit*](https://oll.libertyfund.org/titles/knight-risk-uncertainty-and-profit)
- <i class="fab fa-youtube"></i> Smith, Vernon, L, 2003, ["Constructivist and Ecological Rationality in Economics (Nobel Prize Lecture)](https://www.nobelprize.org/prizes/economic-sciences/2002/smith/lecture/)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("4.5-slides")
```