---
title: "5.1 — Measuring Market Power — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Thursday, April 13, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we begin looking at regulation and antitrust to achieve more competitive markets. We begin with the extremely thorny problems of actually identifying relevant markets and measuring market power.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fas fa-book"></i> Ch. 8.2.2, 12.1.1, 19 in [Churcha nd Ware, 2000](https://works.bepress.com/jeffrey_church/23/)

### Recommended Readings:

- [<i class="fas fa-file-pdf"></i> Bresnahan, 1982, "The Oligopoly Solution Concept is Identified"](/files/readings/Bresnahan-1982.pdf)
- [<i class="fas fa-external-link-square-alt"></i> U.S. Department of Justice, 2010, "Horizontal Merger Guidelines"](https://www.justice.gov/atr/horizontal-merger-guidelines-08192010)
- [<i class="fas fa-external-link-square-alt"></i> U.S. Department of Justice, 2017, "Herfindal Hirschman Index"](https://www.justice.gov/atr/herfindahl-hirschman-index)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("5.1-slides")
```