---
title: "5.3 — Antitrust II: The Paradox — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Tuesday, April 25, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we continue our examination of antitrust law by looking at how antitrust theory and practice have dramatically changed since the 1970s, largely as a result of Robert Bork’s book *The Antitrust Paradox*, and the adoption of the [“Consumer Welfare standard.”](https://en.wikipedia.org/wiki/Consumer_welfare_standard) 

We also turn to take a more critical look at the history of the U.S. in the [Progressive Era](https://en.wikipedia.org/wiki/Progressive_Era) to address three questions:
1. Why were most economists against (or at least, not in favor of) the Sherman Act?
2. Did the industries accused of forming “trusts” and monopolies *actually* behave like monopolies?
3. What was the original intent and purpose of the Sherman Act?

## <i class="fas fa-book-reader fa-lg"></i> Readings

### Recommended Readings:

- [<i class="fas fa-file-pdf"></i> Crandall and Winston, 2003, "Does Antitrust Policy Improve Consumer Welfare? Assessing the Evidence"](/files/readings/Crandall-Winston-2003.pdf)
- [<i class="fas fa-file-pdf"></i> Kovacic and Shapiro, 2000, "Antitrust Policy: A Century of Economic and Legal Thinking"](/files/readings/Kovacic-Shapiro-2000.pdf)
- [<i class="fas fa-file-pdf"></i> DiLorenzo, 1985, "The Origins of Antitrust: An Interest-Group Perspective"](/files/readings/DiLorenzo-1985.pdf)
- [<i class="fas fa-file-pdf"></i> DiLorenzo and High, 1988, "Antitrust and Competition, Historically Considered"](/files/readings/DiLorenzo-High-1988.pdf)
- [<i class="fas fa-file-pdf"></i> McGee, 1958, "Predatory Price Cutting: The Standard Oil (N.J.) Case"](/files/readings/McGee-1958.pdf)
- [<i class="fas fa-file-pdf"></i> Demsetz, 1973, "Industry Structure, Market Rivalry, and Public Policy"](/files/readings/Demsetz-1973.pdf)
- [<i class="fas fa-file-pdf"></i> Bork, 1966, "Legislative Intent and the Policy of the Sherman Act"](/files/readings/Bork-1966.pdf)
- [<i class="fas fa-file-pdf"></i> Hazlett, 1992, "The Legislative History of the Sherman Act Re-Examined"](/files/readings/Hazlett-1992.pdf)
- [<i class="fas fa-file-pdf"></i> Stigler, 1971, "The Theory of Economic Regulation"](/files/readings/Stigler-1971.pdf)
- [<i class="fas fa-file-pdf"></i> Stigler, 1982, "The Economists and the Problem of Monopoly"](/files/readings/Stigler-1982.pdf)
- [<i class="fas fa-file-pdf"></i> Williamson, 1968, "Economies as an Antitrust Defense: The Welfare Tradeoffs"](/files/readings/Williamson-1968.pdf)

- [<i class="fas fa-podcast"></i> *EconTalk*: "Michael Munger on Antitrust"](https://www.econtalk.org/michael-munger-on-antitrust/)
- [<i class="fas fa-podcast"></i> *NPR Planet Money*: "Antitrust 1: Standard Oil"](https://www.npr.org/sections/money/2019/02/15/695131832/antitrust-1-standard-oil)
- [<i class="fas fa-podcast"></i> *NPR Planet Money*: "Antitrust 2: The Paradox"](https://www.npr.org/sections/money/2019/02/20/696342011/antitrust-2-the-paradox)
- [<i class="fas fa-podcast"></i> *NPR Planet Money*: "Antitrust 3: Big Tech"](https://www.npr.org/sections/money/2019/02/22/697170790/antitrust-3-big-tech)
- [<i class="fas fa-podcast"></i> *Capitalisn't*: "Antitrust Pt 1: The Establishment"](https://www.capitalisnt.com/episodes/a2603962)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("5.3-slides")
```