---
title: "5.4 — Antitrust III: The Hipsters — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Thursday, April 27, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we finish our examination of antitrust by looking at the latest debates in antitrust theory and practice, specifically the rise of the [“Neo-Brandeisians”](https://en.wikipedia.org/wiki/New_Brandeis_movement) or (more pejoratively) “Hipster antitrust.” Much of the current discourse about potentially changing antitrust theory and practice revolve around platforms, or “two-sided markets” and the rise of big tech companies like Amazon, Google, and Facebook/Meta. Several antitrust cases are outstanding against these firms, and there are political debates and proposed legislation regarding fundamental changes to antitrust law and enforcement.

## <i class="fas fa-book-reader fa-lg"></i> Readings

### Recommended Readings:

- [<i class="fas fa-file-pdf"></i> Munger, 2022, "Giants Among Us: Do We Need a New Antitrust Paradigm?"](/files/readings/Munger-2022.pdf)
- [<i class="fas fa-podcast"></i> *EconTalk*: "Michael Munger on Antitrust"](https://www.econtalk.org/michael-munger-on-antitrust/)

#### On Path Dependency/Network Economies

- [<i class="fas fa-file-pdf"></i> David, 1985, "Clio and the Economics of QWERTY"](/files/readings/David-1985.pdf)
- [<i class="fas fa-file-pdf"></i> Leibowitz and Margolis, 1990, "The Fable of the Keys"](/files/readings/Leibowitz-Margolis-1990.pdf)
- [<i class="fas fa-external-link-square-alt"></i> BBC: Why we can’t give up this odd way of typing](https://www.bbc.com/worklife/article/20180521-why-we-cant-give-up-this-odd-way-of-typing)
- [<i class="fas fa-external-link-square-alt"></i> Snopes: Are U.S. Railroad Gauges Based on Roman Chariots?](https://www.snopes.com/fact-check/railroad-gauge-chariots/)

#### On Neo-Brandeisian Antitrust

- [<i class="fas fa-file-pdf"></i> Khan, 2017, "Amazon's Antitrust Paradox"](/files/readings/Khan-2017.pdf)
- [<i class="fas fa-file-pdf"></i> Wright, Dorsey, Rybnicek, and Klick, 2018, "Requiem for a Paradox: The Dubious Rise and Inevitable Fall of Hipster Antitrust"](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3249524)
- [<i class="fas fa-file-pdf"></i> Elzinga and Webber, 2017, “Louis Brandeis and Contemporary Antitrust Enforcement”](https://digitalcommons.tourolaw.edu/cgi/viewcontent.cgi?article=2808&context=lawreview)
- [<i class="fas fa-file-pdf"></i> Khan, 2018, “The New Brandeis Movement: America’s Antimonopoly Debate”](https://academic.oup.com/jeclap/article/9/3/131/4915966)

### On Specific Antitrust Cases/History
- [<i class="fas fa-external-link-square-alt"></i> WSJ (Apr 23, 2020): "Amazon Scooped Up Data From Its Own Sellers to Launch Competing Products"](https://www.wsj.com/articles/amazon-scooped-up-data-from-its-own-sellers-to-launch-competing-products-11587650015)
- [<i class="fas fa-podcast"></i> *NPR Planet Money*: "Antitrust 3: Big Tech"](https://www.npr.org/sections/money/2019/02/22/697170790/antitrust-3-big-tech)
- [<i class="fas fa-podcast"></i> *Capitalisn't*: "Antitrust Pt 2: The Populists"](https://www.capitalisnt.com/episodes/bd103ef4)
- [<i class="fas fa-podcast"></i> *Google: The New Vampire Squid? With Dina Srinivasan"](https://www.capitalisnt.com/episodes/should-we-breakup-google-with-dina-srinivasan)
- [<i class="fas fa-podcast"></i> *NPR Planet Money*: "Hey Google, Are You Too Big?"](https://www.npr.org/2020/10/21/926168583/hey-google-are-you-too-big)
- [<i class="fas fa-podcast"></i> *NPR Planet Money*: "The Case Against Facebook"](https://www.npr.org/2020/12/16/947160910/the-case-against-facebook)
- [U.S. Department of Justice, 2020, "Justice Department Sues Monopolist Google For Violating Antitrust Laws"](https://www.justice.gov/opa/pr/justice-department-sues-monopolist-google-violating-antitrust-laws)
- [U.S. Department of Justice, 2023, "Justice Department Sues Google for Monopolizing Digital Advertising Technologies"](https://www.justice.gov/opa/pr/justice-department-sues-google-monopolizing-digital-advertising-technologies)
- [Wikipedia: *Epic Games v. Apple*](https://en.wikipedia.org/wiki/Epic_Games_v._Apple)
- [Wikipedia: Proposed acquisition of Activision Blizzard by Microsoft](https://en.wikipedia.org/wiki/Proposed_acquisition_of_Activision_Blizzard_by_Microsoft)
- [CNBC, 2023: "Meta acquisition of Within reportedly approved by court in loss for FTC"](https://www.cnbc.com/2023/02/01/ftc-loses-attempt-to-block-meta-acquisition-of-within.html)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("5.4-slides")
```