---
title: "5.5 — Reforming Online Data — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Tuesday, May 2, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview


## <i class="fas fa-book-reader fa-lg"></i> Readings

### Required Readings:

- [<i class="fas fa-file-pdf"></i> Posner & Weyl, 2016, "Data as Labor", Chapter 5 in *Radical Markets*](/files/readings/Posner-Weyl-Ch5.pdf)
- [<i class="fas fa-podcast"></i> *EconTalk*: "Glen Weyl on "](https://www.econtalk.org/michael-munger-on-antitrust/)


## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("5.5-slides")
```