---
title: "5.5 — Reforming Online Data — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Tuesday, May 2, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

For this final topic, I want to discuss something I have been mulling over for a long time, but never have really explored sufficiently: how we might reform data management on the internet. This is in part a response to the rise of dominant platforms, privacy concerns, antitrust concerns, and the rise of blockchain/Web3 technology. 

There's a lot worth exploring here, so I recommend you look at both the Weyl and the Munger readings - each comes both as a reading and a podcast option which covers relevant ideas contained in each reading. 

## <i class="fas fa-book-reader fa-lg"></i> Readings

### Required Readings:

- [<i class="fas fa-file-pdf"></i> Posner & Weyl, 2016, "Data as Labor", Chapter 5 in *Radical Markets*](/files/readings/Posner-Weyl-Ch5.pdf)
- [<i class="fas fa-podcast"></i> *Mercatus Center Hayek Program Podcast*: " Glen Weyl on the Myths and Benefits of Automation"](https://podcasts.apple.com/ca/podcast/hayek-program-podcast/id1069881407?i=1000456022970)

- [<i class="fas fa-podcast"></i> *EconTalk*: "Michael Munger on Antitrust"](https://www.econtalk.org/michael-munger-on-antitrust/) (ending starting around 1 hr 6 min mark)
- [<i class="fas fa-external-link-square-alt"></i>  Munger, 2022, “Three Topics for Young Political Economists”](https://www.aier.org/article/three-topics-for-young-political-economists/)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("5.5-slides")
```